### 2020-11-15 latex/acrobot.tex
First editing pass of rotation for acrobot

To prove this intuitive argument rigorously, we define the Poincar\'{e} section
\(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

@@

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
enough \(I\), the Poincar\'{e} section \(P(\mu_0)\) is positive, monotonically
increasing, and \(\lim \limits_{\mu_0 \to \pi} P(\mu_0) = \infty\).

Because \(\dot{\alpha} > 0\), any orbit which begins and remains in \(\mathcal{O}\)
must spiral clockwise around the origin until it intersect the \(q_u\) axis.
Once it does, the Poincar\'{e} argument demonstrates it must spiral away
from the origin; hence, there are no closed orbits and the origin is a repeller.

Once the orbit reaches an intersection point \(\mu_0\) on the \(q_u\)-axis such
that \(P(\mu_0) > \pi\), the orbit will no longer return to the \(q_u\)-axis; it
will escape \(\mathcal{O}\) entirely.
This must always happen in finite time, which implies the orbit must also escape
any compact set in \(mathcal{O}\) in finite time. 
By Definition \ref{defn:energy-gain}, the VNHC is injecting energy into the
acrobot on \(\mathcal{O}\).

If we have \(I < 0\), it is equivalent to using the constraint 
\(q_a = -\bar{q}_a \arctan(I p_u)\). 
By the same analysis, this injects energy on \(mathcal{O}\) in negative-time, so the
constraint dissipates energy on \(mathcal{O}\).

A hybrid supervisor that flips the sign of \(I\) based on the current
value of \(\mu\) will stabilize any energy level of oscillation.
As was the case in the VLP, this supervisor should implement some kind of
hysteresis to avoid infinite switching.
% Empty proof environment to put QED square on the right of the page
\begin{proof}[\unskip\nopunct]
\end{proof}

\subsection*{Perturbation Analysis for Rotations}
Recall that the domain of oscillations for the nominal pendulum of the acrobot
is 
\[
    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
Outside of \(mathcal{O}\), the acrobot is experiencing rotations.
We wish to find a new set of coordinates \((\alpha,\mu)\) and perform a similar
analysis.
During oscillations we showed that the energy of oscillations is determined by
the intersection with the \(q_u\)-axis; 
when the acrobot is rotating, the energy is instead determined by the
intersection on the \(p_u\)-axis.
Hence, we choose \(\mu\) to be this intersection point.
Note that \(mathcal{O}\) intersects the \(p_u\) axis when 
\(\mu^2 = 60m^2 g l^3\), so we must have 
\(\mu > \sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u > 0\) 
or \(\mu < -\sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u < 0\).
The energy level set associated with \(\mu\) is
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\mu^2}{10ml^2}\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:rotation-pu2}
    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\mu^2}{10 ml^2}
    .
\end{equation}
On this level set, \(q_u\) ranges fully between \(]-pi,\pi]\), so our angle of
rotation is uniquely parameterized by \(\alpha = q_u\).
Since \(\mu\) does not change sign along the rotation, we have the smooth
relationship
\begin{align}\label{eqn:rotation-Tinv}
    q_u &= \alpha
    , \\
    p_u &= \sign{\mu}\sqrt{\mu^2 - 30 m^2 g l^3\left(1 - c_\alpha \right)}
    .
\end{align}

We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\), with
inverse 
\begin{align}\label{eqn:rotation-T}
    \alpha &= q_u
    , \\
    \mu &= \sign{p_u}\sqrt{p_u^2 + 30 m^2 g l^3 \left(1 - c_u \right)}
    .
\end{align}

From this point on, we do the exact same procedure as with oscillations.
MATLAB symbolic operations evaluate the nominal system's dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\alpha}_n &=  \sign(\mu_0) 
    \frac{\sqrt{\mu^2 - 30m^2gl^3(1 - c_\alpha)}}{5ml^2}
    .
\end{align}
As is expected, \(\dot{\alpha}_n\) does not change sign because the acrobot is
rotating;
if \(\mu_0 > 0\), the orbit traces a curve going from \(\alpha = -\pi\) to 
\(\alpha = \pi\), and if \(\mu_0 < 0\) the curve goes in the reverse direction.
For small enough \(I > 0\) we must also have 
\(\dot{\alpha} = \sign{\mu_0}\) and can hence use \(\alpha\) as a new time
variable.

Take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) and initial
state \(\mu_0\).
We expand the solution of this into the same form as
~\eqref{eqn:acrobot-muhat-approx} where once again we have
\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
The dynamics for \(\hat{\mu}_1\) are
\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
    \dot{\hat{\mu}}_1 =
    \frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
        - c_\alpha\mu_0^2
    \right)}{
    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
    }
    =: b(\alpha,\mu_0)
    .
\end{equation}
The solution to ~\eqref{eqn:acrobot-rot-mu1-dot} is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) = \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma
    .
\]
Since one full rotation amounts to \(\alpha\) traversing \(2\pi\)rad in a clockwise
direction (\ie \(\alpha\) going from \(0\) to \(\sign{\mu_0}2\pi\)), 
we define a Poincar\'{e} section as the point reached on the \(p_u\) axis after
every rotation:
\[
    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0)
    .
\]
If \(|P(\mu_0)| > |\mu_0|\), then the rotation has
driven the orbit further from the origin.
In other words, \(|p_u|\) has increased across the rotation.
If this is true for any \(\mu_0\), the acrobot is gaining energy.

The function \(b(\alpha,\mu_0)\) is even in both its inputs, and is 
\(2\pi\)-periodic in \(\alpha\).
Hence, we have 
\begin{align*}
    \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0) &=
    \int \limits_0^{\sign{\mu_0}2\pi} b(\sigma,\mu_0)d\sigma
    , \\
     &=\sign{\mu_0}\int\limits_0^{2\pi}b(\sigma,\mu_0)d\sigma
     , \\
     &= \sign{\mu_0} \hat{\mu}_1(2\pi,\mu_0)
     .
\end{align*}
Furthermore, \(\hat{\mu}_1(2\pi,\mu_0) = \hat{\mu}_1(2\pi,|\mu_0|)\).
This implies that
\[
    |P(\mu_0)| = \left|\sign{\mu_0}\left(
        |\mu_0| + I \hat{\mu}_1(2\pi,|\mu_0|)\right)
    \right|
    = P(|\mu_0|)
    .
\]
Therefore, if \(P(|\mu_0|) > |\mu_0|\), then the acrobot is gaining energy.
Showing that \(\hat{\mu}_1(2\pi,\mu_0) > 0\) for \(\mu_0 > \sqrt{60m^2gl^3}\) is
sufficient for this condition to hold.

Unlike in the case of oscillations, we cannot adimensionalize
\(b(\alpha,\mu_0)\) and prove that this condition will be true for all acrobots. 
Let \(U = ]\sqrt{60m^2gl^3},\bar{\mu}[\), and assume that 
\(\hat{\mu}_1(2\pi,\mu_0) > 0 \) for all \(\mu_0 \in U\).
If this is true, then by perturbation analysis we can claim that there exists a
small enough \(I\) such that the Poincar\'{e} section \(P(\mu_0)\) is monotonic
on \(U\).

Define the set of points in \(\Gamma\) which have pseudo-radius in \(U\) by
\[
    W := \left\{(q_u,p_u) \in \Gamma \mid |\mu(q_u,p_u)| \in U \right\}
    ,
\]
as in Figure \ref{fig:acrobot-rot-W}.
Monotonicity of \(P(\mu_0)\) means there are no closed orbits of 
\((q_u,p_u)\) on \(W\). 
Furthermore, \(|p_u|\) is increasing every time the orbit hits the
\(p_u\)-axis.

\begin{figure}
    \centering
    % TODO: Display the set W, which is the two rotation bands at the top and
    % bottom with outer limit +- \bar{mu}  and inner limit +- sqrt{60m^2gl^3}.
    % Show also that the inner limit connects to D
    %\includestandalone[width=0.5\textwidth]{images/acrobot_rot_W}
    \caption{The set \(W\) of points which have pseudo-radius 
        \(|\mu| \in ]\sqrt{60m^2gl^3}, \bar{\mu}[\). The inner boundary of \(W\)
        is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
        oscillation domain \(mathcal{O}\).}
    \label{fig:acrobot-rot-W}
\end{figure}

We conclude that \(|\mu(q_u,p_u)|\) will eventually reach \(\bar{\mu}\). 
We cannot claim this fits the definition  of energy injection 
(since \((0,0) \notin W\)), but we can still say that \((q_u,p_u)\) will
eventually depart \(W\) through the outer boundary in finite time.
The acrobot is thereby gaining energy in the literal sense that its momentum
is increasing on \(W\).

If we set \(I < 0\), we have \(P(|\mu_0|) < |\mu_0|\).
Hence, \(|p_u|\) is decreasing every time the orbit hits the \(p_u\)-axis, so
the acrobot will eventually depart \(W\) through the inner boundary. 
The acrobot is losing momentum, and therefore it is losing energy on \(W\).

\subsection*{Connecting Oscillations and Rotations}
To complete this proof, we will prove the acrobot gains energy on the set 
\[
    E_{\bar{\mu}} := 
    \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\mu}) \right\}
    .
\]
% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
% therefore exit D and enter W.

>>>>

To prove this intuitive argument rigorously, we define the Poincar\'{e} section
\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

@@

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
enough \(I\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive,
monotonically increasing, and 
\(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).

% Empty proof environment to put QED square on the right of the page
\begin{proof}[\unskip\nopunct]
\end{proof}

\subsection*{Perturbation Analysis for Rotations}
The rotation region for the nominal pendulum is the set
\[
    \mathcal{R} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) > E(\pi,0)\right\}
    .
\]
We wish to find a new set of coordinates \((\beta,\rho)\) for the acrobot
and perform a similar analysis to what we did for oscillations.
In the oscillation region \(\mathcal{O}\) we showed that the nominal pendulum's
energy (and hence, its orbit) is uniquely determined by the intersection point
\(\mu\) with the \(q_u\)-axis; 
when the nominal pendulum is rotating, the energy is instead determined by the
intersection \(\rho\) on the \(p_u\)-axis.
Note that the boundary of \(\mathcal{O}\) intersects the \(p_u\) axis when 
\(p_u^2 = 60m^2 g l^3\); 
hence, for a rotation we must have \(\rho > \sqrt{60 m^2 g l^3}\) if the
orbit has momentum \(p_u > 0\) 
or \(\rho < -\sqrt{60 m^2 g l^3}\) if the orbit has momentum \(p_u < 0\).
The energy level set associated with \(\rho\) is
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:rotation-pu2}
    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
    .
\end{equation}
On this level set, \(q_u\) takes all values on \(\mathbb{S}^1\), so our angle of
rotation is uniquely parameterized by \(\beta = q_u\).
Since \(\rho\) does not change sign along the rotation, we have the smooth
relationship
\begin{align}\label{eqn:rotation-Tinv}
    q_u &= \beta
    , \\
    p_u &= \sign{\rho}\sqrt{\rho^2 - 30 m^2 g l^3\left(1 - c_\beta \right)}
    .
\end{align}

Inverting this relationship gives our (smooth) transformation
\begin{equation*}
    T : \mathbb{S}^1 \times \R \rightarrow 
    \mathbb{S}^1 \times ]-\infty,-60m^2gl^3[\, \cup \, ]60m^2gl^3,\infty[ 
    ,
\end{equation*}
where \(T(q_u,p_u) = (\beta,\rho)\) is defined by
\begin{align}\label{eqn:rotation-T}
    \beta &= q_u
    , \\
    \rho &= \sign{p_u}\sqrt{p_u^2 + 30 m^2 g l^3 \left(1 - c_u \right)}
    .
\end{align}

Computing the acrobot's constrained dynamics in \((\beta,\rho)\)-coordinates and
setting \(I = 0\) yields the dynamics of the nominal pendulum.
MATLAB evaluates those dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\rho}_n &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\beta}_n &=  \sign{\rho} 
    \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
    .
\end{align}
As expected, \(\dot{\beta}_n\) does not change sign on \(\mathcal{R}\) because
a rotating pendulum does not change direction:
if \(\rho_n > 0\), the orbit traces a curve going from \(\beta_n = -\pi\) to 
\(\beta_n = \pi\); if \(\rho_n < 0\) the curve goes in the reverse direction.
By continuity of the acrobot's constrained dynamics 
\((\dot{\beta},\dot{\rho})\), we must also have 
\(\sign{\dot{\beta}} = \sign{\rho_0}\) for small enough \(I\).
Hence, we can use \(\beta\) as a new ``time" variable on \(\mathcal{R}\) 
by performing a time transformation \(t = \tau(\beta)\).
This produces the time-scaled system \(\hat{\rho}(\beta) = \rho(\tau(\beta))\)
with dynamics
\[
    \begin{cases}
        \dot{\hat{\rho}} = \frac{\dot{\rho}}{\dot{\beta}}
        , \\
        \hat{\rho}(0) = \rho_0
        .
    \end{cases}
\]
For a perturbation approach, be expand the solution of these dynamics into
\begin{equation}\label{eqn:acrobot-rhohat-approx}
    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) + I
    \hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
    ,
\end{equation}
where \(R(\beta,\rho_0,I)\) is \(O(I^2)\).

From ~\eqref{eqn:khalil-perturbation-nominal} we know
\(\hat{\rho}_0\) is the solution to the nominal system
\[
    \begin{cases} 
        \dot{\hat{\rho}}_0 = 0
        , \\
        \hat{\mu}_0(0) = \rho_0
        ,
    \end{cases}
\]
which means \(\hat{\mu}_0(\beta,\rho_0) \equiv \rho_0\) for all \(\beta\).
Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} 
gives us the following dynamics for \(\hat{\rho}_1\):
\begin{equation}\label{eqn:acrobot-rho1-dot}
  \begin{cases}
    \dot{\hat{\rho}}_1 =
    \frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
        - c_\beta\rho_0^2
    \right)}{
    |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
    }
    =: b(\beta,\rho_0)
     , \\
     \hat{\rho}_1(0) = 0
     .
 \end{cases}
\end{equation}
The solution to ~\eqref{eqn:acrobot-rho1-dot} is given by
\[
    \hat{\rho}_1(\beta,\rho_0) = \int \limits_0^\beta b(\sigma,\rho_0)d\sigma
    .
\]

Supposing we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\),
then one full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
We define a Poincar\'{e} section as the point reached on the \(p_u\) axis after
every rotation:
\[
    P(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
      + R(2\pi,I)
    .
\]
If \(|P(\rho_0)| > |\rho_0|\), then the rotation has
driven the orbit further from the origin.
In other words, \(|p_u|\) has increased across the rotation.
If this is true for any \(\rho_0\), the acrobot is gaining momentum, which in
turn means it is gaining energy.

The function \(b(\beta,\rho_0)\) is even in both its inputs, and is 
\(2\pi\)-periodic in \(\beta\).
Hence, we have 
\begin{align*}
    \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0) &=
    \int \limits_0^{\sign{\rho_0}2\pi} b(\sigma,\rho_0)d\sigma
    , \\
     &=\sign{\rho_0}\int\limits_0^{2\pi}b(\sigma,\rho_0)d\sigma
     , \\
     &= \sign{\rho_0} \hat{\rho}_1(2\pi,\rho_0)
     .
\end{align*}
Furthermore, \(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_1(2\pi,|\rho_0|)\).
This implies that
\[
    |P(\rho_0)| = \left|\sign{\rho_0}\left(
        |\rho_0| + I \hat{\rho}_1(2\pi,|\rho_0|)\right)
    \right|
    = P(|\rho_0|)
    .
\]
Hence, if \(P(|\rho_0|) > |\rho_0|\), then a single rotation starting at \(\rho_0\)
will drive the acrobot's orbit further from the origin.
Equivalently, one need simply guarantee that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
for all \(\rho_0 > \sqrt{60m^2gl^3}\).

Unfortunately, we cannot adimensionalize
\(b(\beta,\rho_0)\) like we did for oscillations. 
Hence, we let \(U = ]\sqrt{60m^2gl^3},\bar{\rho}[\) and assume that 
\(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all \(\rho_0 \in U\).
We conclude that, for small enough \(I\), the Poincar\'{e} section 
\(P(\rho_0)\) is monotonically diverging from 0 on \(U\)
and the acrobot is gaining momentum.

\subsection*{Energy Gain}
We will prove the acrobot gains energy on the set 
\[
    E_{\bar{\mu}} := 
    \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    .
\]
% Can we use the jacobian? The spectrum is hard to find.

Recall that on \(\mathcal{O}\) we have \(\dot{\alpha} > 0\). 
Hence, any orbit which begins and remains in \(\mathcal{O}\)
must spiral clockwise around the origin until it intersect the \(q_u\) axis.
Once it does, the Poincar\'{e} section \(P_\mathcal{O}(\cdot)\) 
demonstrates the orbit must spiral away from the origin; 
this implies there cannot be any closed orbits on \(\mathcal{O}\), and that the
origin is a repeller.

Once the orbit reaches an intersection point \(\mu_0\) on the \(q_u\)-axis such
that \(P_\mathcal{O}(\mu_0) > \pi\), the orbit will no longer return to the
\(q_u\)-axis. 
Instead, it will escape \(\mathcal{O}\) entirely.
This must always happen in a finite number of iterations of the 
Poincar\'{e} section, which means the orbit escapes \(\mathcal{O}\) in finite
time., which implies the orbit must also escape
any compact set in \(\mathcal{O}\) in finite time. 
By Definition \ref{defn:energy-gain}, the VNHC is injecting energy into the
acrobot on \(\mathcal{O}\).

If we have \(I < 0\), it is equivalent to using the constraint 
\(q_a = -\bar{q}_a \arctan(I p_u)\). 
By the same analysis, this injects energy on \(\mathcal{O}\) in negative-time, so the
constraint dissipates energy on \(\mathcal{O}\).

A hybrid supervisor that flips the sign of \(I\) based on the current
value of \(\mu\) will stabilize any energy level of oscillation.
As was the case in the VLP, this supervisor should implement some kind of
hysteresis to avoid infinite switching.

% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
% therefore exit D and enter W.

Define the set of points in \(\Gamma\) which have pseudo-radius in \(U\) by
\[
    W := \left\{(q_u,p_u) \in \Gamma \mid |\rho(q_u,p_u)| \in U \right\}
    ,
\]
as in Figure \ref{fig:acrobot-rot-W}.
Monotonicity of \(P(\rho_0)\) means there are no closed orbits of 
\((q_u,p_u)\) on \(W\). 
Furthermore, \(|p_u|\) is increasing every time the orbit hits the
\(p_u\)-axis.

\begin{figure}
    \centering
    % TODO: Display the set W, which is the two rotation bands at the top and
    % bottom with outer limit +- \bar{mu}  and inner limit +- sqrt{60m^2gl^3}.
    % Show also that the inner limit connects to D
    %\includestandalone[width=0.5\textwidth]{images/acrobot_rot_W}
    \caption{The set \(W\) of points which have pseudo-radius 
        \(|\rho| \in ]\sqrt{60m^2gl^3}, \bar{\rho}[\). The inner boundary of \(W\)
        is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
        oscillation domain \(mathcal{O}\).}
    \label{fig:acrobot-rot-W}
\end{figure}

We conclude that \(|\rho(q_u,p_u)|\) will eventually reach \(\bar{\rho}\). 
We cannot claim this fits the definition  of energy injection 
(since \((0,0) \notin W\)), but we can still say that \((q_u,p_u)\) will
eventually depart \(W\) through the outer boundary in finite time.
The acrobot is thereby gaining energy in the literal sense that its momentum
is increasing on \(W\).

If we set \(I < 0\), we have \(P(|\rho_0|) < |\rho_0|\).
Hence, \(|p_u|\) is decreasing every time the orbit hits the \(p_u\)-axis, so
the acrobot will eventually depart \(W\) through the inner boundary. 
The acrobot is losing momentum, and therefore it is losing energy on \(W\).
